---
title: "A Sociology of the Family"
format: 
  revealjs:
    controls: true
    theme: [default, custom.scss]
---

```{r warning = FALSE, error = FALSE, message = FALSE}
# Loading libraries
library(tidyverse)
library(gghighlight)

# Define color palette
c_palette <- c("#20c997",
               "#3498db",
               "#fd7e14",
               "#e74c3c")
```


# Defining Families {.theme-section}  
What makes a family?  
  

::: notes
Q&A with students -- live together? legal ties? pets?
:::

------------------------------------------------------------------------

<iframe width="100%" height="100%" src="https://www.youtube.com/embed/yaeiCEro0iU" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen>

</iframe>

::: notes
11 mins.
:::

## Definitions  

::: {.incremental .highlight-last}
-   Groups of people who are related by genetics, marriage, or choice, and who share material, emotional, and economic resources (CrashCourse)
-   **Groups of related people bound by connections that are biological, legal, or emotional** (Cohen 2020)
:::

  
::: notes
-   There is not a single definition of family.
-   Different criteria are used for different reasons (e.g., biological, legal, emotional).
-   Different people and countries have different definitions.
-   Critique CrashCourse definition -- have to share resources?
-   This is the simplest, working definition used by Philip Cohen.
:::

## Personal family

-   People to whom we feel related and who we expect to define us as members of their family as well\
-   Families of choice

::: notes
-   Choice: marriage; close friends -- fictive kin
:::

## Legal family

-   A group of individuals related by birth, marriage, or adoption

::: notes
-   No universal legal definitions, and legal implications vary
-   State-imposed legal definition (state refers to government at all levels)
-   Implications for the distribution of resources
-   Definition seems straightforward, but its meaning may be contested.
-   All the words in the definition can be contested.
-   Most contested word is marriage (marriage equality)
:::

## Households  

-   A group of people that lives and eats separately from other groups

::: notes
-   households are sometimes families, families are sometimes households
:::

## Households cont.  

```{r}  

# Data: https://www150.statcan.gc.ca/n1/daily-quotidien/220713/g-a004-eng.htm
type <- c("One-census family", "Other-census family", "2 + non-census family", "1 person")
year <- c("2001", "2001", "2001", "2001",
          "2006", "2006", "2006", "2006",
          "2011", "2011", "2011", "2011",
          "2016", "2016", "2016", "2016",
          "2021", "2021", "2021", "2021")
prop <- c(64.6, 5.9, 3.7, 25.7,
          63.6, 5.9, 3.7, 26.8,
          62.0, 6.3, 4.1, 27.6,
          61.2, 6.5, 4.1, 28.2,
          59.6, 6.6, 4.4, 29.3)
data <- data.frame(type, year, prop)

data$type <- factor(data$type, 
                    levels = c("One-census family", "Other-census family", 
                               "2 + non-census family", "1 person"), ordered = FALSE)

p <- ggplot(data, aes(x = factor(year), y = prop, fill = type, label = prop)) +
  geom_col(position = position_stack(reverse = TRUE),
           width = .5) +
  # Set the position to its complementary
  geom_text(position = position_stack(reverse = TRUE, vjust = .5),
            show.legend = FALSE, color = "white") +
  coord_flip() +
  scale_x_discrete(limits=rev) +
  theme_minimal(base_size = 20) +
  theme(legend.position="top",
        panel.grid.minor = element_blank()) +
  scale_fill_manual(values = c_palette) +
  ggtitle("Distribution (%) of household types in Canada") +
  labs(x        = NULL, 
       y        = NULL,
       fill     = NULL,
       legend   = NULL,
       caption  = "Source: Census of Population 2001 to 2021 (3901)") 

p + gghighlight(type == "One-census family", label_key = type, use_group_by = FALSE)

```

::: notes
Census of Population 2001 to 2021 (3901)
## https://www150.statcan.gc.ca/n1/daily-quotidien/220713/g-a004-eng.htm
:::

## Households cont.  
```{r}
p + gghighlight(type == "One-census family" |type == "Other-census family", 
                label_key = type, use_group_by = FALSE)
```


## Households cont.  
```{r}
p 
```

# Theories of Families {.theme-section}  
  
## Institutional Arena

A framework identifying a space where interactions occur

1.  Family arena

2.  State arena

3.  Market arena

::: notes
The goods and services available in the market, combined with the services and policies of the government, affect decisions individuals make about family life.
:::

## Institutional Arena cont. {.smaller}

|                          | Family                                                                    | State                                                   | Market                                                                |
|------------------|------------------|------------------|-------------------|
| **Type of Interaction**  | people practice intimacy, childbearing and socialization, and caring work | behavior is legally regulated through laws and policies | labor for pay, economic exchange, and wealth accumulation takes place |
| **Organizational Units** | Families                                                                  | Legislatures and agencies                               | Companies                                                             |
| **Individual Roles**     | Family members                                                            | Citizens and residents                                  | Workers, owners, and consumers                                        |

::: notes
Talk through examples of each.\
\* state -- marriage license\
\* market -- childcare
:::

## Broad Perspectives

**Consensus**\
society as the collective expression of shared norms and values

. . .

**Conflict**\
focus on power relations and inequality

::: notes
**consensus:** stability -- breadwinner/homemaker were complementary and in harmony\
macro -- order and stability for society\
micro -- emotional and economic support\
**conflict:** change -- focus on the competing interests of family members to understand family issues.\
macro -- inequality may be between the family system and the work/employment/economic system\
micro -- inequality may occur between certain family members
:::

## Structural functionalism  
family as a social institution that performs essential functions for society and ensures its stability (Mitchell 2021)  

::: notes  
in line with the consensus perspective  
- "the family" is universal (breadwinner/homemaker)
:::

## Exchange  
individuals or groups with different resources, strengths, and weaknesses enter into mutual relationships to maximize their own gains  
  
::: notes  
in line with the consensus perspective  
Individuals cannot provide for all of their own needs on their own, so they must exchange with others  
Economist Gary Becker -- positive interpretation about the unequal partnerships between men and women  
breadwinner-homemaker family - an employed father, a mother without employment, and their children  
the husband and wife perform separate and different roles, and this “exchange” benefits the marriage and the family  
division of housework is a major subject  
family wage: the amount necessary for a male earner to provide for family  
criticized because it presumes that resources between men and women are equal  
:::
  
## Feminism  
seeks to understand and ultimately reduce inequality between men and women

::: notes  
in line with the conflict perspective  
opposed to structural functionalism  
Gender inequality is central to family life  
The family arena is where gender roles are learned and formed  
family structure is socially constructed (and unequal)  
Critical of the breadwinner-homemaker family model  
:::

## Symbolic Interaction  
ability of humans to see themselves through the eyes of others  

::: notes  
enact social roles based on others’ expectations  
studying social change but is not strictly coming from the conflict perspective  
Meaning can only be understood by studying the actions (behaviors) of individuals  
individuals see themselves through the eyes of others and act according to the expectations of others  
:::


## Demography  
family behavior and household structures contribute to larger population processes

::: notes  
the study of populations  
focuses on birth, death, and migration  
must look at these things in context, hence an emphasis on the timing of life events  
:::

## Life Course  
family trajectories of individuals and groups as they progress through their lives, in social and historical context  

::: notes  
progression from childhood to adulthood, looking for patterns and what is typical  
Traditional “normal” progression used to include marriage and then parenthood  
Sequences of family events now more diverse and complicated  
Goal is to place family events in historical context  
Cohort: a group of people who experience an event together at the same point in time  
:::
