---
title: "Code-along 01"
author: "Your Name Goes Here"
editor: visual
execute:
  error: true
format: pdf
---

## Packages

We'll use the following packages:

-   `here()` (relative file paths)
-   `tidyverse()` (data wrangling)
-   `gssr()` (U.S. General Social Survey data)
-   `gssrdoc()` (GSS documentation)

## Install `here()` and `tidyverse()`

Let's first install the two packages that are available on CRAN.

Copy and paste the following code into your Console pane. Then hit
enter.

`install.packages("here")`

Then, do the same to install the tidyverse package.

`install.packages("tidyverse")`

## Install `gssr()` and `gssrdoc()`

Install `gssr` from `ropensci` universe. Copy and paste the following
code into your Console pane. Then hit enter.

`install.packages('gssr', repos = c('https://kjhealy.r-universe.dev', 'https://cloud.r-project.org'))`

Also install `gssrdoc.`Copy and paste the following code into your
Console pane. Then hit enter.

`install.packages('gssrdoc', repos = c('https://kjhealy.r-universe.dev', 'https://cloud.r-project.org'))`

## Load the packages

```{r}
#| label: load

library(here)
library(tidyverse)
library(gssr)
library(gssrdoc)
```

## Environment

```{r}
#| label: enviro

# software documentation
sessionInfo()

```

## Project structure

Let's set up your project structure using the `here()` package.

First, let's establish our project directory.

```{r}
#| label: here

# set the file path to the root of the project
here()

```

Next, we'll create folders within our project using `here()` and
`dir.create()`

```{r}
#| label: base-folders

# Create base folders
dir.create(here("data"), recursive = TRUE)
dir.create(here("code-alongs"), recursive = TRUE)
dir.create(here("milestones"), recursive = TRUE)
dir.create(here("project"), recursive = TRUE)
```

Now, we'll create sub-folders using `here()` and `dir.create()`

```{r}
#| label: sub-folders

# Create project sub-folders
dir.create(here("project", "data"), recursive = TRUE)
dir.create(here("project", "scripts"), recursive = TRUE)
dir.create(here("project", "outputs"), recursive = TRUE)

```

Check your work by reporting a list of folders and or files in the
R-project folders and sub-folder.

```{r}
#| label: list-files

# Your SOC6302 class folder
list.files(path = here())

# Your "Project" sub-folder
list.files(path = here("project"))

```

## Save code-along

Save this code-along in your newly created "code-along" sub-folder.

There's no command in the R console to save scripts or Quarto files —
you use the editor's [File]{style="color: #3498DB"} \> [Save
As]{style="color: #3498DB"} or Ctrl+S.

## Meet your data

We're going to use data from the [U.S. General Social Survey
(GSS)](https://gss.norc.org/us/en/gss/about-the-gss.html).

Let's load your data.

```{r}
#| label: gssdata

# Load the data (will appear in your Global Environment pane)
data(gss_all)

# Preview the datatable which is automatically named gss_all
gss_all

```

A “tibble” is another name for “tidy dataset,” meaning that the data is
organized in structured, clear rows and columns. “(75,699 × 6,867)”
means the dataset contains 75,699 rows and 6,867 columns. In social
sciences, rows are commonly referred to as “observations” and columns as
“variables.” In our case, there are 75,699 observations (e.g.,
respondents) and 6,867 variables.

You can also load the GSS data for a specific survey year.

```{r}
#| label: gss24

# Get the data only for the 2024 survey respondents
gss24 <- gss_get_yr(2024)

# look at the first 6 rows of the dataframe
head(gss24)

```

With your mouse, go to the environment panel (upper-right) and click on
the `gss24` object. It pops up and you can browse through it.

This is often a good idea to get a first feel for the data, but only if
your dataset is relatively small.

The [GSS
documentation](https://gss.norc.org/us/en/gss/get-documentation.html) is
also available online in .pdf form.

The .pdfs will be useful for general overviews.

For specific variable information, it will be helpful to use the
documentation you'll load into RStudio.

```{r}
#| label: gsscodebook

# Load the codebook
data(gss_dict)

```

To see the variables available in the dataset, use the `names()`
command. This command is best to use with smaller datasets.

```{r}
#| label: names
#| eval: false # copy/paste this code in the Console; don't run in Quarto

names(gss_all)
```

## Variables

For information about a specific GSS variable, type `?varname` at the
console.

In the output pane, the Help tab will show the variable documentation.

Example:

```{r}
#| label: meovrwrk-help
#| eval: false # copy/paste this code in the Console; don't run in Quarto

?meovrwrk
```

**Notes:**

Variable name: meovrwrk\
Variable label: Men hurt family when focus on work too much\
1994 was the first year of the survey.\
695 respondents agreed with the statement.\
iap -- missing. Values: the numeric and response category key (1 =
strongly agree)

We often want to know which years a question was asked.

We can find this out for one or more variables with `gss_which_years()`.

```{r}
#| label: which-years

gss_which_years(gss_all, meovrwrk)

```

If run in the console, to see all rows, wrap the code in the `print()`
command:

`print(gss_which_years(gss_all, meovrwrk), n = 40)`

You can access the variables (i.e., columns) using the `$` operator, as
shown using the table() function

(NOTE: The variable names are case sensitive. In this dataset, all
variables are lowercase.)

```{r}
#| label: meovrwrk-table-gss_all

table(gss_all$meovrwrk)

```

2436 respondents were coded as 1 on this variable. What does that mean?
(Look at the help page.)

Let's look at only the 2024 respondents. Change the code to show just
the `gss24` respondents. Then, add text to your code-along that
interprets the results for the `2` value.

```{r}
#| label: meovrwrk-table-gss24

# Replace gss_all with gss24 and then run the code.
table(gss_all$meovrwrk)
```

## Quarto: Render

Finally, let's `render` your code-along-01 and view the results.
