---
title: "SOC214H"
---

```{r, include=FALSE}

if(!require(tidyverse)){
  install.packages("tidyverse")
  library(tidyverse)
}

if(!require(knitr)){
  install.packages("knitr")
  library(knitr)
}

if(!require(kableExtra)){
  install.packages("kableExtra")
  library(kableExtra)
}

if(!require(googledrive)){
  install.packages("googledrive")
  library(googledrive)
}

if(!require(googlesheets4)){
  install.packages("googlesheets4")
  library(googlesheets4)
}

if(!require(tinytex)){
  install.packages("tinytex")
  library(tinytex)
}

## sheets reauth with specified token and email address
drive_deauth()
drive_auth(
  cache = ".secrets",
  email = "joannapepin@gmail.com"
)

# Give google sheets the same authorization
gs4_auth(token = drive_token())

```

# Sociology of the Family

Winter 2023

**Instructor:** Joanna R. Pepin

**Office Hours:** Virtual, by appointment

**Email:** j.pepin\@utoronto.ca

**Teaching Assistants:** TBD

# Course Description

This course provides a critical analysis of families using a life course perspective. We will pay attention to historical transformations of families in society and to family transitions over the life course of individuals. This course emphasizes the linked lives of individuals to each other, to institutions such as work, and to their cultural contexts. The intersections between the family and other social institutions, as well as growing diversity between families will be central foundations of this course.

We will examine trends in family life such as racial and economic inequality, technology's transformation of romantic relationships, and demographic trends. Other topics include historical trends in courtship, family formation, parenting, families and work, and divorce and remarriage.

## Learning Outcomes

The course will introduce you to sociological concepts and issues pertaining to contemporary family sociology. You will finish the course with a better understanding of the differences between families and the ways that social forces shape families. You will apply the sociological perspective to the institution of the family and examine how social factors such as race, class, and gender shape contemporary families. By the end of this course, students should be able to:

-   Understand and evaluate the family as a socially and culturally constructed institution.
-   Think critically about how race, class, gender, age, and social location shape how individuals experience family life within society.
-   Have an awareness of historical and demographic trends shape and change family patterns in American society.
-   Be able to form arguments and think critically about family life and broader society.

## Prerequisite

The prerequisite to take this course is **SOC100H1**. As per university guidelines, those without this requirement will be removed at time of discovery and without notice.

# Required Texts

There is no textbook for this course. See Quercus for required readings available electronically.

# Schedule

```{r message =FALSE, echo=FALSE}
sched <- read_sheet("10NtiaxfX8Mj61gsIwV4vSWBQ9WTY2lv746YkLMi6o3Y", sheet = "Schedule")
tests <- c(5, 11, 14)

```

```{r, echo = FALSE}
options(knitr.kable.NA = '')
kable(sched) %>%
  kable_styling(bootstrap_options = "striped", position = "center") %>%
  row_spec(0,     color = "white",   background = "#18BC9C") %>%
  row_spec(7,     color = "#18BC9C", bold = T) %>%
  row_spec(tests, color = "#E74C3C", bold = T)


```

# Evaluation

## Components

All assignments, evaluations, and tests must be submitted via Quercus.

```{r echo = FALSE}
Output <- c("Flipped Quizzes", "Test 1", "Test 2", "Fact Check Paper")
Proportion <- c("25%", "25%", "25%", "25%")
summary <- data.frame(Output, Proportion)
```

```{r echo = FALSE}
kable(summary) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left") %>%
  row_spec(0, color = "white", background = "#18BC9C")
```

### Flipped Quizzes

You will create 5 multiple choice questions (5% each) related to the course material (readings, lectures, videos, activities, etc.) and share them with your classmates to assist in study review. Detailed instructions about developing appropriate quiz questions will be provided.

Two quizzes are due by Test 1, two quizzes are due by Test 2, and one quiz is due by the end of the term (same due date as the final paper). Quiz content may only be used from modules that correspond with the upcoming exam (or final paper). For example, you can't use content from Week 2 after Test 1 has already passed or content from week 8 after Test 2.

### Tests

There will be 2 tests over the course of the semester during the class times noted on the schedule. Tests will be open book and will not be cumulative but all materials in this course (i.e., readings, lectures, videos, and discussions) are considered "fair game."

Tests will be available on Quercus and will be a mix of multiple choice and short answer questions. You will have two hours to complete each test.

### Fact Check Paper

You will be asked to write a 500-1500 word examination of a claim you find in the media related to a course topic. Think of this as a writing assignment structured as: "I read that XXXX, but in reality the research says YYYY."

Additional assignment details will be provided later in the course.

## Arts & Sciences Grading Scale

```{r echo = FALSE}
Percentage <- c("90-100", "85-89", "80-84", 
                "77-79",  "73-76", "70-72", 
                "67-69",  "63-66", "60-62", 
                "57-59",  "53-56", "50-52", "0-49")
Letter     <- c("A+", "A", "A-", 
                "B+", "B", "B-", 
                "C+", "C", "C-", 
                "D+", "D", "D-", "F")
GPA        <- c("4.0", "4.0", "3.7", 
                "3.3", "3.0", "2.7",
                "2.3", "2.0", "1.7",
                "1.3", "1.0", "0.7", "0.0")
Definition <- c("Excellent", " ",  " ",
                "Good",      " ",  " ",
                "Adequate",  " ",  " ",
                "Marginal",  " ",  " ", "Inadequate")
grading <- data.frame(Percentage, Letter, GPA, Definition)
```

```{r echo = FALSE}
kable(grading) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left") %>%
  row_spec(0, color = "white", background = "#18BC9C")
```
