---
title: "SOC395H"
---

```{r, include=FALSE}

if(!require(tidyverse)){
  install.packages("tidyverse")
  library(tidyverse)
}

if(!require(knitr)){
  install.packages("knitr")
  library(knitr)
}

if(!require(kableExtra)){
  install.packages("kableExtra")
  library(kableExtra)
}

if(!require(googledrive)){
  install.packages("googledrive")
  library(googledrive)
}

if(!require(googlesheets4)){
  install.packages("googlesheets4")
  library(googlesheets4)
}

if(!require(tinytex)){
  install.packages("tinytex")
  library(tinytex)
}

## sheets reauth with specified token and email address
drive_deauth()
drive_auth(
  cache = ".secrets",
  email = "joannapepin@gmail.com"
)

# Give google sheets the same authorization
gs4_auth(token = drive_token())

```

# Intimate Relationships

Winter 2023

**Instructor:** Joanna R. Pepin

**Office Hours:** Virtual, by appointment

**Email:** j.pepin\@utoronto.ca

**Teaching Assistants:** TBD

# Course Description

This course provides an in-depth examination of the development and maintenance of intimate relationships across the life course, with an emphasis on sociological theoretical perspectives, contextual factors, and key sociological policy debates.

## Learning Outcomes

## Prerequisite

# Required Texts

# Schedule

# Evaluation

## Components

All assignments, evaluations, and tests must be submitted via Quercus.

### Critical Reflections
These 1 page reports involve engaging critically with specified week’s readings and making connections between them to the real world. You will need to select five modules to submit reports prior to the week's discussion.

### Interview Project
The major assignment for this course is your interview project, which has three parts. Early in the semester, students will be interviewing someone about their thoughts on an aspect of intimate relationships. You will then write up a summary and preliminary analysis of the interview (4-5 pages) and post it on Quercus so classmates can analyze the data for their final reports (8 pages). Detailed requirements and rubrics to be posted.

### Policy Essay
Students will select one family policy or one recent family court case found in a country of their choice. Examples of family policies include the provision of childcare services, of maternity/parental leaves, or of financial child benefits. Family court cases could focus on issues such as the rights and obligations of gay couples, the number of parents a child can have, or the rights and obligations of separated parents in terms of custody and child support. The project report should first briefly present the policy or court decision, describe the pursued goal or intent and situate it in the broader context of the country. Secondly, the report should take a critical stance and discuss, for example, counter arguments to the ruling or some unexpected and perhaps adverse effects of the policy. Finally, the report should conclude regarding the benefits/disadvantages of the policy or ruling and point to remedial measures, if needed.

The project report should be 4-5 pages long (double space, Times New Roman font 12, and one-inch margins), excluding references and other supportive material. Students need to submit the choice of their topic and country for approval by the Professor no later than ...

### Data Design
What don't we know about an aspect of intimate relationships? Situate your research question in the literature.Cite 4-6 peer reviewed articles. What are your hypotheses? What quantitative data would you need to collect to answer the question? Design your study.

## Arts & Sciences Grading Scale

```{r echo = FALSE}
Percentage <- c("90-100", "85-89", "80-84", 
                "77-79",  "73-76", "70-72", 
                "67-69",  "63-66", "60-62", 
                "57-59",  "53-56", "50-52", "0-49")
Letter     <- c("A+", "A", "A-", 
                "B+", "B", "B-", 
                "C+", "C", "C-", 
                "D+", "D", "D-", "F")
GPA        <- c("4.0", "4.0", "3.7", 
                "3.3", "3.0", "2.7",
                "2.3", "2.0", "1.7",
                "1.3", "1.0", "0.7", "0.0")
Definition <- c("Excellent", " ",  " ",
                "Good",      " ",  " ",
                "Adequate",  " ",  " ",
                "Marginal",  " ",  " ", "Inadequate")
grading <- data.frame(Percentage, Letter, GPA, Definition)
```

```{r echo = FALSE}
kable(grading) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left") %>%
  row_spec(0, color = "white", background = "#18BC9C")
```
